### 2019-02-01 abstract.tex
moving to new repo

The involvement of manual operations and domain expertise makes traditional model optimization process expensive, time-consuming and not guaranteed the optimal solution. The emergence of serverless computing offers the potential to optimize models in an auto-scaling and micro-billing fashion. In this paper, we present Seneca -- a framework for machine learning model optimization on serverless computing platform. Its automated pipeline consists of packaging, deployment, serverless function optimization and hyperparameter tuning, which provides a cost-effective, high-performing and reliable model optimization service. We evaluate Seneca by performance, cost and latency analysis on multiple regression and classification applications. The result demonstrates Seneca streamlines model deployment and optimization in agile and economical fashion.