### 2019-02-02 intro.tex

The scale and elasticity of cloud computing systems have 
fueled remarkable innovation and unprecedented commercial 
investment.  Cloud users ``rent'' virtualized resources (while sharing
the underlying physical resources) on a pay-per-use basis
in exchange for availability guarantees specified via service level
agreements (SLAs). Uniquely, cloud systems can be configured 
to add and remove resources and services,
automatically and on-demand, based on load 
or as required by executing applications.

To date however, cloud use has been limited to large-scale 
implementations of enterprise services, such as
object storage, databases, queuing systems, and web and application servers.
The reason for this is that it is very
challenging, even for expert users, to configure complex
distributed systems for application use, 
and to leverage the elasticity (auto-scaling) that
clouds offer. To address this challenge, cloud providers 
have started to offer cloud programming and execution environments
under the moniker of \textit{serverless} 
computing~\cite{xxx,yyy,zzz}.  Serverless
systems automatically configure, manage, and scale applications
to significantly simplify and broaden participation in (by less expert
users) the use of cloud systems.
%xxx: http://shivaram.org/publications/pywren-socc17.pdf
%yyy: http://cidrdb.org/cidr2019/papers/p119-hellerstein-cidr19.pdf
%zzz: L. Wang, M. Li, Y. Zhang, T. Ristenpart, and M. Swift. “Peek- ing Behind the Curtains of Serverless Platforms”. In: USENIX ATC. Boston, MA: USENIX Association, 2018

Using the serverless model, applications developers upload 
arbitrary computations in high level languages
as stateless functions to these cloud-hosted services, along with a
specification conditions under which each
function should be triggered.  Functions are invoked (triggered)
automatically by the cloud in response to updates to other cloud services
(e.g. storage, queues, notification services, among others). They
communicate, persist, and access data only 
via their inputs or shared, remote storage.  
As a result, serverless applications are inherently elastic and can 
implement highly concurrent and parallel tasks.
In public clouds, users pay a small fee for the resources their 
functions use during execution, resulting in very low cost cloud use.
Although now available from all public cloud providers and as open 
source for private cloud systems, 
AWS Lambda~\cite{awslambda} was the first and is now the 
most popular and most advanced serverless public cloud service 
available today.
%awslambda: https://docs.aws.amazon.com/lambda/latest/dg/lambda-dg.pdf

In this work, we investigate the efficacy of using AWS Lambda
for tuning machine learning applications.
To date, Lambda has not be heavily used for training and evaluating
machine learning models because of a concern that 
doing so will result in high overhead (i.e. be costly) because
of the stateless nature of serverless functions~\cite{ref:onesteptwostep}.
At the same time, identifying the ``best'' configuration for advanced
machine learning models is challenging given the large number of configuration
options, i.e., hyperparameters, typical of models today.
Hyperparameters governs the learning process of machine learning applications.
Given that parameter sweeps are embarrassingly parallel, we believe
that such tuning is a good fit for the serverless model.
To investigate this potential, its overhead, and to simplify the 
use of Lambda for training, testing, and evaluation machine learning models, 
we design and develop a new system and toolset called \textit{Seneca}.

Specifically, Seneca implements, packages, and deploys 
machine learning applications as stateless functions to AWS Lambda.
It then orchestrates exhaustive evaluation of hyperparameter settings
to identify the best performing model (for a given dataset) by
comparing the error and accuracy across models.  Users present
Seneca with their application, a range of values for 
each hyperparameter (or the default can be used), and a dataset.
Seneca produces, tests, and evaluates models for all combinations 
of hyperparameters and returns to the user, 
the set of parameters (or the model itself) that produces
the best cross-validation score (e.g. lowest squared error or highest
accuracy).

We deploy Seneca on AWS Lambda and evaluate its tuning performance, cost,
and memory use for six machine learning applications and datasets. We 
find that Seneca is effective for quickly training these
models and identifying the best performing hyperparameter setting for 
these applications. We also find 
Seneca to be very low cost and able to automatically 
identify the 
best memory configuration for each application further lowering its 
by 10-30\%.
We intend to make Seneca and the applications available as open source
when/if this paper is accepted for publication.  In the sections
that follow, we overview our design and implementation of Seneca and then
present our empirical methodology and results.

\ignore{
Recent work~\cite{ref:onesteptwostep} argues however that AWS Lambda

Serverless~\cite{ref:serverless} computing is a rising function-based (also known as Functions-as-a-Service~\cite{ref:faas}) programming and deployment paradigm on top of cloud infrastructure, in which programmers manifest business logic without concerning provisioning servers by modular functions that are triggered by incoming events and return computation results in serialized format like XML or JSON. Such events include HTTP requests from clients, possibly proxied by API Gateway, or communications between cloud services based on heterogeneous protocols. Depending on the supported runtimes from cloud vendors, functions could be written in various high-level languages (i.e. NodeJS, Python, Java, etc.) and take advantage of serverless SDK to interact with other cloud services. Until recently, the leading serverless platform, Amazon Web Services(AWS) Lambda, expands language support by custom runtime~\cite{ref:customruntime} where software engineers can implement functions by any programming languages.

Primarily, there are four evolutionary advantages that serverless architecture brings to cloud computing. First and foremost, it raises the abstraction layer from virtual machines to function containers~\cite{ref:container} that abstracts away the responsibilities of infrastructure management from DevOps~\cite{ref:devops} personnel to the cloud provider, in terms of scaling, patching, provisioning, error-handling and terminating underlying bare metal servers or clusters. Such shift allows more agile development life cycle and more product-oriented programming iteration. Secondly, the event-driven and paralleled container runtime of serverless function provides more fine-grained computational resource isolation and usage. Hence, serverless architecture makes cloud operation more secure and energy-efficient compared to traditional virtual machine instance runtime. Thirdly, serverless architecture decomposes monolithic application code base even further from microservices~\cite{ref:microservices} into granular functions, which are easier to be created, maintained and refactored by team or individual programmer. Furthermore, Function-as-a-Service applications could be autoscaled independently with respect to modular functions depending on the volume of request. This capability also contributes to the energy efficiency and eliminating throttling bottlenecks. Lastly, serverless computing provides "Pay-per-execution" pricing model~\cite{ref:pricing}. In a real world case study~\cite{ref:serverless_econ}, the server operational cost on serverless platform could be reduced dramatically, given a relatively lower amount of Transactions Per Second (TPS). According to these merits, the stateless, event-driven and embarrassingly parallel tasks are potentially cost-effective with lower latency by running on serverless framework. 
}