### 2019-02-02 conclusion.tex

In this paper, we present Seneca, a new framework for simplifying and expediting
the training and testing of machine learning models in terms of their prediction 
power using very low cost cloud computing services.
Specifically, Seneca leverages the AWS Lambda serverless cloud service for autoscaled
(elastic) and parallel execution of hyperparameter search and selection for machine learning
applications.  
Users provide Seneca with the application code and libraries, 1+ datasets, and the list 
of possible hyperparameter settings.  Seneca uses this 
information to automatically construct an AWS Lambda function that
that invokes the application, imports and splits the dataset into training
and testing datasets, and constructs, tests, and evaluates (i.e. scores) a machine
learning model for a particular configuration of hyperparameter settings.  
Seneca orchestrates parallel execution of these functions for all possible 
combinations of hyperparameter values specified and returns the best scoring model
and configuration to the user (for future use on other datasets).

We present the design, implementation, and cost optimization for Seneca.
The optimization automaticallly optimizes function memory use to reduce the cost
of AWS Lambda use.  Our empirical evaluation using
multiple applications for regression and classification (using
the models for prediction and not explanitory value), shows that Seneca is able to quickly
identify the best performing model for the applications and datasets that we consider.  
We also show that Seneca memory optimization reduces the cost of using
Seneca by 10-30\% for the applications studied.