### 2019-02-03 conclusion.tex

We present a new framework, called Seneca, for simplifying and expediting
the training and testing of machine learning models
using low cost cloud services.
Specifically, Seneca leverages the AWS Lambda for autoscaled
(elastic) and parallel execution of hyperparameter search 
and selection for machine learning models.  
Users provide Seneca with the application code and libraries, 1+ datasets, and the list 
of possible hyperparameter settings.  
%Seneca uses this 
%information to automatically construct models using parallel functions in AWS Lambda. 
%an AWS Lambda function that
%that invokes the application, imports and splits the dataset into training
%and testing datasets, and constructs, tests, and evaluates (i.e. scores) a machine
%learning model for a particular configuration of hyperparameter settings.  
Seneca uses this information to orchestrate parallel execution of these functions for all possible 
combinations of hyperparameter values specified and returns the best scoring model
and configuration to the user (for future use on other datasets).

We present the design, implementation, and cost optimization for Seneca.
The optimizer automatically optimizes function memory use to reduce the cost
of AWS Lambda use.  Our empirical evaluation using
multiple applications for regression and classification (using
the models for prediction and not explanatory value), shows that Seneca is able to quickly
identify the best performing model for the applications and datasets that we consider.  
We also show that Seneca memory optimization reduces the cost of using
Seneca by 10-30\% for the applications studied.