### 2019-02-26 conclusion.tex
chandras pass

We present a new framework, called Seneca, for simplifying and expediting
the training and testing of machine learning models
using low cost cloud services.
Specifically, Seneca leverages the AWS Lambda for autoscaled
(elastic) and parallel execution of hyperparameter search 
and selection for machine learning models.  
Users provide Seneca with the application code and 
libraries, 1+ datasets, and the list 
of possible hyperparameter settings.  
Seneca uses this information to automatically
configure and deploy these functions in parallel for all possible 
combinations of hyperparameter values specified. Seneca
returns the best scoring model
and configuration to the user (for future use on other datasets).

We present the design, implementation, and cost optimization for Seneca.
The optimizer automatically optimizes function memory use to reduce the cost
of AWS Lambda use.  Our empirical evaluation using
multiple applications for regression and classification,
shows that Seneca is able to quickly
identify the best performing hyperparameter setting 
for the applications and datasets that we consider.  
We also show that Seneca memory optimization reduces the cost of using
Seneca by 10-30\% for the applications studied.