### 2019-02-26 related_work.tex

We next overview related work. We discuss contributions that focus
on serverless computing, automatic packaging, 
and machine learning (ML) model optimization.

Multiple researchers have investigated the efficacy and overhead of the
serverless programming model and 
implementations~\cite{ref:jonas2017occupy,ref:onesteptwostep,ref:baldini2017,ref:lin2018tracking}.
%The authors in \cite{ref:onesteptwostep} evaluate the cost and performance of using
%serverless for data-intensive computing.  
They identify some weaknesses in AWS
Lambda for data-intensive computing, including the inability to 
effectively train models.
Our work shows that it is possible and cost effective to use
AWS Lambda to train models for the subset of ML applications that we 
consider.

PyWren~\cite{ref:jonas2017occupy} shows that stateless functions
are a natural fit for distributed computing models. The technique
abstracts away cluster management overhead and is ideal for 
embarrassingly parallel jobs.  ExCamera~\cite{ref:encoding} presents a 
framework for running general-purpose parallel tasks (encoding 4K video) 
on a commercial serverless platform using multithreading. 
Cirrus~\cite{ref:jonas2019cloud} attempts to train ML models 
using a parameter server and serverless functions. 
An empirical comparison of resource utilization and performance isolation
is provided in~\cite{ref:peeking} for AWS Lambda, 
Azure Functions, and Google Cloud Functions.

The serverless framework~\cite{ref:serverless_framework} provides
automated packaging and deployment for serverless functions across clouds. 
The serverless framework uses CloudFormation\footnote{\url{https://aws.amazon.com/cloudformation/}} for deployment in AWS Lambda, which introduces additional cost. 
The cloud infrastructure provisioning 
framework Terraform\footnote{\url{https://www.terraform.io/}}
also provides automated deployment of functions to serverless platforms
(however, without replicating the Lambda runtim, introducing deployment errors).
Seneca uses a local docker container to avoid this cost and overhead. 
%In addition, AWS CodeDeploy~\cite{ref:codedeploy} serves as component in the continuous integration and delivery pipeline for serverless architectures, but it also introduces monetary cost overhead to do so.

Automated hyperparameter optimization is the focus of
many projects. Google Vizier~\cite{ref:vizier} provides a service for 
black-box optimization. Optunity~\cite{ref:claesen2014hyperparameter} 
and Hyperopt\footnote{\url{http://hyperopt.github.io/hyperopt/}}
provide a Python library for hyperparameter tuning. 
%This can be considered as precursory work of Hyperopt~\cite{ref:hyperopt}, 
%which is a more advanced Python library for serial and parallel hyperparameter 
%optimization over awkward search space. 
Hyperas\footnote{\url{http://maxpumperla.com/hyperas/}} adds another 
abstraction layer to hyperopt to facilitate hyperparameter tuning for 
Keras~\footnote{\url{https://keras.io/}}. However, no extant work of which we are aware leverages
serverless to perform hyperparameter tuning and memory optimization in parallel 
for ML applications.